### 2016-11-29 real/draft.tex
[draft] move bold symbols into separate file
[draft] Include Hans's changes, rewrite basic formulae

\newcommand{\hmod}  {{\mathcal{H}}}  % Model
\newcommand{\ldef}{\;{:}{=}\;}
\newcommand{\rdef}{\;{=}{:}\;}
\newcommand{\realnumbers}{\mathbb{R}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\cond}{\,|\,}
\newcommand{\NBD}{\mathrm{NB}}
\newcommand{\var}{\mathrm{var}}

\newcommand{\smA}{{\scriptscriptstyle A}}
\newcommand{\smB}{{\scriptscriptstyle B}}
\newcommand{\smH}{{\scriptscriptstyle H}}
\newcommand{\smK}{{\scriptscriptstyle K}}
\newcommand{\smL}{{\scriptscriptstyle L}}
\newcommand{\smM}{{\scriptscriptstyle M}}
\newcommand{\smN}{{\scriptscriptstyle N}}
\newcommand{\smR}{{\scriptscriptstyle R}}

\newcommand{\smo}{{\scriptscriptstyle 1}}
\newcommand{\smt}{{\scriptscriptstyle 2}}
\newcommand{\smr}{{\scriptscriptstyle 3}}
\newcommand{\smf}{{\scriptscriptstyle 4}}
\newcommand{\smv}{{\scriptscriptstyle 5}}
\newcommand{\smx}{{\scriptscriptstyle 6}}
\newcommand{\sms}{{\scriptscriptstyle 7}}
\newcommand{\sme}{{\scriptscriptstyle 8}}
\newcommand{\smn}{{\scriptscriptstyle 9}}

\newcommand{\bma}{{\bm{a}}}
\newcommand{\bmb}{{\bm{b}}}
\newcommand{\bmc}{{\bm{c}}}
\newcommand{\bmd}{{\bm{d}}}
\newcommand{\bme}{{\bm{e}}}
\newcommand{\bmf}{{\bm{f}}}
\newcommand{\bmg}{{\bm{g}}}
\newcommand{\bmh}{{\bm{h}}}
\newcommand{\bmi}{{\bm{i}}}
\newcommand{\bmj}{{\bm{j}}}
\newcommand{\bmk}{{\bm{k}}}
\newcommand{\bml}{{\bm{\ell}}}
\newcommand{\bmm}{{\bm{m}}}
\newcommand{\bmn}{{\bm{n}}}
\newcommand{\bmo}{{\bm{o}}}
\newcommand{\bmp}{{\bm{p}}}
\newcommand{\bmq}{{\bm{q}}}
\newcommand{\bmr}{{\bm{r}}}
\newcommand{\bms}{{\bm{s}}}
\newcommand{\bmt}{{\bm{t}}}
\newcommand{\bmu}{{\bm{u}}}
\newcommand{\bmv}{{\bm{v}}}
\newcommand{\bmw}{{\bm{w}}}
\newcommand{\bmx}{{{\bm{x}}}}
\newcommand{\bmy}{{\bm{y}}}
\newcommand{\bmz}{{\bm{z}}}
\newcommand{\bmA}{{\bm{A}}}
\newcommand{\bmB}{{\bm{B}}}
\newcommand{\bmC}{{\bm{C}}}
\newcommand{\bmD}{{\bm{D}}}
\newcommand{\bmE}{{\bm{E}}}
\newcommand{\bmF}{{\bm{F}}}
\newcommand{\bmG}{{\bm{G}}}
\newcommand{\bmH}{{\bm{H}}}
\newcommand{\bmI}{{\bm{I}}}
\newcommand{\bmJ}{{\bm{J}}}
\newcommand{\bmK}{{\bm{K}}}
\newcommand{\bmL}{{\bm{L}}}
\newcommand{\bmM}{{\bm{M}}}
\newcommand{\bmN}{{\bm{N}}}
\newcommand{\bmO}{{\bm{O}}}
\newcommand{\bmP}{{\bm{P}}}
\newcommand{\bmQ}{{\bm{Q}}}
\newcommand{\bmR}{{\bm{R}}}
\newcommand{\bmS}{{\bm{S}}}
\newcommand{\bmT}{{\bm{T}}}
\newcommand{\bmU}{{\bm{U}}}
\newcommand{\bmV}{{\bm{V}}}
\newcommand{\bmW}{{\bm{W}}}
\newcommand{\bmX}{{\bm{X}}}
\newcommand{\bmY}{{\bm{Y}}}
\newcommand{\bmZ}{{\bm{Z}}}

\newcommand{\bmalpha}{{\bm{\alpha}}}
\newcommand{\bmbeta}{{\bm{\beta}}}
\newcommand{\bmchi}{{\bm{\chi}}}
\newcommand{\bmdelta}{{\bm{\delta}}}
\newcommand{\bmepsilon}{{\bm{\varepsilon}}}
\newcommand{\bmphi}{{\bm{\phi}}}
\newcommand{\bmgamma}{{\bm{\gamma}}}
\newcommand{\bmeta}{{\bm{\eta}}}
\newcommand{\bmiota}{{\bm{\iota}}}
\newcommand{\bmkappa}{{\bm{\kappa}}}
\newcommand{\bmlambda}{{\bm{\lambda}}}
\newcommand{\bmmu}{{\bm{\mu}}}
\newcommand{\bmnu}{{\bm{\nu}}}
\newcommand{\bmomega}{{\bm{\omega}}}
\newcommand{\bmpi}{{\bm{\pi}}}
\newcommand{\bmpsi}{{\bm{\psi}}}
\newcommand{\bmrho}{{\bm{\rho}}}
\newcommand{\bmsigma}{{\bm{\sigma}}}
\newcommand{\bmtau}{{\bm{\tau}}}
\newcommand{\bmtheta}{{\bm{\theta}}}
\newcommand{\bmupsilon}{{\bm{\upsilon}}}
\newcommand{\bmxi}{{\bm{\xi}}}
\newcommand{\bmzeta}{{\bm{\zeta}}}

@@

\DeclareMathOperator{\Expect}{\mathbb{E}}
\newcommand{\expect}[1]{\Expect\left[#1\right]}
\DeclareMathOperator{\GammaDist}{Gamma}
\DeclareMathOperator{\GaussianDist}{\mathcal{N}}
\newcommand{\Kalpha}{{K_\alpha}}
\newcommand{\Kbeta}{{K_\beta}}
\newcommand{\npack}{n_p}
\newcommand{\lmax}{\ell_{\rm max}}
\newcommand{\Lmax}{L_{\rm max}}
\newcommand{\Lumtot}{Q}
\newcommand{\lumtot}{q}
\newcommand{\Lum}{L}
\newcommand{\lum}{\ell}
\newcommand{\rmdx}[1]{\mbox{d} #1 \,} % differential
\newcommand{\firstDeriv}[1]{\frac{\partial}{\partial #1}}
\newcommand{\secDeriv}[1]{\frac{\partial^2}{\partial #1^2}} % second partial derivative
\newcommand{\secPartial}[2]{\frac{\partial^2}{\partial #1 \, \partial #2}} % second partial derivative
\newcommand{\tardis}{TARDIS}

@@

\section{Introduction} \label{sec:intro}

\begin{itemize}
  \item radiative transfer
  \item get uncertainty from single run
  \item cover case when few packets in a bin
\end{itemize}

@@

@@

We consider the total number of packets $\npack$ fixed, for example
chosen by a user. In contrast, the number of packets in a bin, $n_b$
is a realization of a random variable $N_b$ which we assume to follow
a Poisson distribution. With this assumption, we neglect bin-by-bin
correlations and treat each bin independently which allows us to focus
on a single bin and to drop the subscript $b$ in the remainder. Hence
\begin{align}
  \lleq{caa}
  p(N\cond\lambda) &= \frac{e^{-\lambda}\lambda^N}{N!}
  \qquad N = 0,1,\ldots,\infty ,
\end{align}
where the dependency of the Poisson mean, $\lambda>0$, on $\nu$ is
implicit. Let $\bmL$ denote the collection of luminosities of packets
in a certain bin. The total luminosity in that bin is then given by
\begin{align}
  \lleq{cab}
  \Lumtot = \sum_{j=1}^{N} \Lum_j\,.
\end{align}
We assume that each luminosity $\Lum_j$, and we usually omit the subscript $j$ to refer to a single luminosity, has a distribution depending on $\nu$ and
other parameters $\bmphi$, and that each packet is logically
independent. Since $N$ follows a Poisson distribution, $\Lumtot$
follows a \emph{compound Poisson process}. The joint distribution then is
\begin{align}
  \lleq{cac}
  (N,\bmL) &\ \sim\ p(N,\bmL\cond \bmphi,\lambda, \nu)
  = p(\bmL\cond N,\bmphi, \nu)\; p(N\cond \lambda) = \prod_{j=1}^N p(\Lum_j\cond \bmphi, \nu) \; p(N\cond \lambda),
\end{align}
where we used that the Poisson term does not depend on $\nu$
explicitly. In general, the distribution of $\Lum_j$ could be a of a
different functional form in each bin. For simplicity, we assume it is
identical but may depend on parameters that differ from bin to bin. In
practice, the distribution of packets usually does not change abruptly
with frequency. We therefore assume that any parameter needed to fully
specify $p(\Lum_j)$ is given explicitly as a continuous function of $\nu$
and the unknown parameters $\bmphi$. This allows us to use a small
number of parameters to describe $p(\Lumtot)$ across the entire frequency
range instead of having a few parameters per bin which could lead to
tens of thousands of parameters for the applications we have in
mind. So we determine $\bmphi$ from fitting all simulated packets
simultaneously whereas $\lambda$ is a per-bin quantity.
\fred{Reformulate when asymptotic approx. included: it has thousands of parameters}

@@

\subsection{Predicting the number of packets in a bin}
\fred{Show reference Poisson analysis here of before?}

We use Bayes' theorem to find the posterior distribution for $\lambda$
given data $n$ and the Poisson hypothesis
\begin{align}
  \lleq{pmd}
  p(\lambda\cond n) %
  &= \frac{p(n\cond\lambda)\,p(\lambda)} {p(n)}
  \ =\ \frac{p(n\cond\lambda)\,p(\lambda)} %
  {\int \rmdx{\lambda} p(n\cond\lambda)\,p(\lambda)}
\end{align}
to find the posterior predictive of $N$ given $n$ taking into account
all possible values of $\lambda$,
\begin{align}
  \lleq{pme}
  p(N\cond n)
  &= \int \rmdx{\lambda}p(N,\lambda\cond n)
  \ =\ \int \rmdx{\lambda} p(N\cond\lambda, n)\,p(\lambda\cond n)
  \ =\ \frac{\int \rmdx{\lambda} p(N\cond\lambda)\,p(n\cond\lambda)\,p(\lambda)}
  {\int \rmdx{\lambda} p(n\cond\lambda)\,p(\lambda)}
\end{align}
where $p(N\cond\lambda,n) = p(N\cond\lambda)$ is independent of $n$.
Common uninformative priors are
\begin{align}
  \lleq{pmf}
  p(\lambda) = c\lambda^{-a},
\end{align}
where $a=0$ gives the uniform prior\footnote{For a uniform prior, the
  case $a{=}0$, we require of course a maximum $\lambda_{\rm max}$. As
  long as $\lambda_{\rm max}$ is large enough, the integral can
  effectively be taken to infinity.}, $a=1/2$ gives the Jeffreys and
reference prior, and $a=1$ is the transformation-group result
advocated by Jaynes\fredmargin{book 12.4.2}. All choices lead to an
improper prior but the posterior predictive is proper provided that $n
\ge 1$; i.e., at least one packet is observed in the bin. The standard
results for the evidence and the posterior predictive are
\begin{align}
  \lleq{pmg}
  p(n) &= %
  \int_0^\infty d\lambda\, p(n\cond\lambda)\,p(\lambda)
  \ =\
  c \int_0^\infty d\lambda\, \frac{e^{-\lambda}\,\lambda^{n-a}}{n!}
  \ =\ c\frac{(n-a)!}{n!} \, ,
  \\
  \lleq{pmh}
  p(N\cond n) %
  &= \frac{n!}{(n-a)!} \int_0^\infty d\lambda\,
  \frac{e^{-2\lambda} \lambda^{N+n-a}}{N!\, n!}
  \ =\ \frac{(N+n-a)!}{N!\,(n-a)!\,2^{N+n+1-a}} \, .
\end{align}

>>>>

\input{symbols}

\newcommand{\hypo}  {{\mathcal{H}}}  % Model
\newcommand{\ldef}{\;{:}{=}\;}
\newcommand{\rdef}{\;{=}{:}\;}
\newcommand{\realnumbers}{\mathbb{R}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\cond}{\,|\,}
\newcommand{\NBD}{\mathrm{NB}}
\newcommand{\var}{\mathrm{var}}

@@

\DeclareMathOperator{\Expect}{\mathbb{E}}
\newcommand{\expect}[1]{\Expect\left[#1\right]}
\newcommand{\expectest}[1]{\widehat{\Expect\left[#1\right]}}
\DeclareMathOperator{\GammaDist}{Gamma}
\DeclareMathOperator{\GaussianDist}{\mathcal{N}}
\DeclareMathOperator{\InvGammaDist}{InvGamma}
\newcommand{\Kalpha}{{K_\alpha}}
\newcommand{\Kbeta}{{K_\beta}}
\newcommand{\npack}{{n_p}}
\newcommand{\lmax}{\ell_{\rm max}}
\newcommand{\Lmax}{{L_{\rm max}}}
\newcommand{\Lumtot}{Q}
\newcommand{\lumtot}{q}
\newcommand{\Lum}{L}
\newcommand{\lum}{\ell}
\newcommand{\rmdx}[1]{\mbox{d} #1 \,} % differential
\newcommand{\firstDeriv}[1]{\frac{\partial}{\partial #1}}
\newcommand{\secDeriv}[1]{\frac{\partial^2}{\partial #1^2}} % second partial derivative
\newcommand{\secPartial}[2]{\frac{\partial^2}{\partial #1 \, \partial #2}} % second partial derivative
\newcommand{\tardis}{TARDIS}
\DeclareMathOperator{\Variance}{\mathbb{V}}
\newcommand{\variance}[1]{\Variance\left[#1\right]}

@@

\section{Introduction (Wolfgang)}

\subsection{Astrophysical context}

\textbf{Define here?}
\begin{enumerate}
\item Basic notation for packets, luminosity as used below
\end{enumerate}

\textbf{Issues that need to be resolved}
\begin{enumerate}
\item Write everything in terms of frequency $\nu$ or wavelength $\lambda$
\end{enumerate}

\subsection*{Main points of this paper}

\begin{enumerate}
\item The first goal of this paper is to show that the Bayesian
  approach to this problem is well-founded and robust.

\item Luminosity data generated by a single simulation provides a
  basis for predicting its uncertainty, i.e.\ how much this luminosity
  can be expected to vary from simulation to simulation.

\item The uncertainty in the luminosity calculated in the Bayesian
  approach is larger than anticipated because it takes into account
  not only the variability of the luminosity per packet, but also the
  distribution of the number of packets.

\item Traditional methods using conventional counting ratio
  (``relative frequency'') can also make predictions for such
  uncertainties, but only if the number of packets is large. By
  contrast, the Bayesian method presented here remains valid also for
  small packet multiplicities.

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Bayesian prediction}

We first sketch the generic notation and relations for Bayesian
probability theory.
%
A particular experiment or simulation generates \textit{data} $D$
which will often come in the form of a list of $n$ measurements,
$D = \bmy = (y_1,y_2,\ldots,y_n)$. In order to understand or explain
this data, the observer constructs one or more \textit{models} or
\textit{hypotheses} $\hypo$. Central to this construction is the
specification of a \textit{likelihood} $p(\bmy\cond \bmtheta,\hypo)$
where $\bmtheta = (\theta_1,\theta_2,\ldots,\theta_K)$ is the set of
unknown parameters specified by the model. If and when the individual
measurements $y_i$ are independent, the joint likelihood is the
product
\begin{align}
  p(\bmy\cond \bmtheta) &= \prod_{i=1}^n p(y_i\cond \bmtheta).
\end{align}
At this level of analysis, the goal
is to find which values for these parameters provide a satisfactory
description of the data. In the Bayesian context, these parameter
values are best described in terms of the \textit{posterior}, which
is related to the likelihood by
\begin{align}
  \lleq{bth}
  p(\bmtheta\cond \bmy)
  &= \frac{p(\bmy\cond\bmtheta)\;p(\bmtheta)}
    {p(\bmy)}
    %= \frac{p(\bmy\cond\bmtheta)\;p(\bmtheta)}
    %{\int d\bmtheta'\, p(\bmy\cond\bmtheta')\;p(\bmtheta')}
\end{align}
where $p(\bmtheta)$, the so-called \textit{prior}, is the probability
assigned to the values of the parameters before any data was taken
into account and the \textit{evidence}
\begin{align}
  p(\bmy) &= \int d\bmtheta'\, p(\bmy\cond\bmtheta')\;p(\bmtheta'),
\end{align}
is fully determined by the likelihood and prior.

The scope of this paper is limited to the relationship between
simulated data generated by radiative transfer simulation codes.
Likewise, all parameters introduced and calculated relate strictly to
model descriptions of the simulation.
%
While Bayesian concepts and methods can be applied to actual
experimental data and related physics parameters, we shall not do so
here.

\hanslong{
Note that if the prior is uniform, it is independent of $\theta$ and
will cancel,
\begin{align}
  \lleq{bthu}
  p(\theta\cond \bmy)
  &= \frac{p(\bmy\cond\theta)} {\int d\theta'\, p(\bmy\cond\theta')}.
\end{align}
}

\hanslong{Bayesian prediction?}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{TARDIS}
\fred{I don't think TARDIS is a good title, too specific}

We consider a computer simulation which, based on tunable input
parameters $\bmtheta$, generates $\npack$ photon packets, each with a
particular frequency $\nu_i$ and luminosity $\ell_i$. Together, these
frequencies and luminosities form the simulation data
$\bmy = \{ (\nu_i,\ell_i): i = 1,\ldots,\npack\}$.  We divide up the
frequency spectrum into bins
$\Delta_b = [\nu_b {-} \tfrac{1}{2} \Delta\nu \,,\, \nu_b {+}
\tfrac{1}{2} \Delta\nu], b = 1,2,\ldots, B$ of width $\Delta\nu$
centered on frequencies $\nu_b$. Let $n_b$ be the number of packets
having a frequency falling into a particular bin $b$, and let $\bml_b$
be the set of luminosities of those $n_b$ packets. The total
luminosity in bin $b$ is the sum of luminosities of those packets
whose frequency falls into $b$,
\begin{align}
  \ell_b = \sum_{i=1}^\npack \bm{1}_{\Delta_b}(\nu_i)\,\ell_i
\end{align}
where the indicator function
$\bm{1}_{[a,b]}(x) = 1$ when $x$ is in the interval
$[a,b]$ and 0 otherwise.

@@

\fred{Merge this section with previous section}

@@

We consider the total number of packets $\npack$ fixed, for example
chosen by a user. In contrast, the number of packets in a bin, $n_b$
is a realization of a random variable $N_b$ which we assume to follow
a Poisson distribution. With this assumption, we neglect bin-by-bin
correlations and treat each bin independently. From now on, we focus
on a single bin and drop the subscript $b$. Hence
\begin{align}
  \lleq{caa}
  p(N\cond\lambda) &= \frac{e^{-\lambda}\lambda^N}{N!}
  \qquad N = 0,1,\ldots,\infty ,
\end{align}
where the dependency of the Poisson mean, $\lambda>0$, on the bin, or
respectively the frequency $\nu$, is implicit. Let $\bmL = (\Lum_1,\dotsc,\Lum_N)$ denote the
collection of luminosities of packets in the bin, then the total
luminosity in that bin is
% and its realization based on the data $\bmy$ are
\begin{align}
  \lleq{cab}
  \Lumtot &= \sum_{j=1}^{N} \Lum_j%, \qquad \lumtot = \sum_{j=1}^{n} \lum_j
  \,.
\end{align}
The distribution of the luminosity for each packet,
$p(\Lum \cond \bmphi)$, depends on parameters $\bmphi$ and potentially
on the frequency but we suppress the latter dependency.  The packets
are logically independent. Since $N$ follows a Poisson distribution,
$\Lumtot$ follows a \emph{compound Poisson process}.
%
The key quantity of interest is the distribution of $\Lumtot$ given
the data $\bmy$, $p(\Lumtot \cond \bmy)$ for the compound Poisson
process. In the following, we rewrite $p(\Lumtot \cond \bmy)$ to
obtain an expression that we can readily evaluate numerically. Using
the law of total probability,
\begin{align}
  \lleq{qaa}
  p(\Lumtot \cond \bmy) &= \int \rmdx{\bmphi} p(\bmphi \cond \bmy) p(\Lumtot \cond \bmphi, \bmy) \,,
\end{align}
where the first term in the integral is the posterior $p(\bmphi \cond \bmy)$ after observing $\bmy$. We can decompose the second term as
\begin{align}
  \lleq{qab}
  p(\Lumtot \cond \bmphi, \bmy) &= \sum_{N=0}^\infty  p(N \cond \bmphi, \bmy) p(\Lumtot \cond N, \bmphi, \bmy)\,,
\end{align}
where $p(N \cond \bmphi, \bmy) = p(N \cond \lambda)$ is given by
\refeq{caa} and the second term in the sum is
\begin{align}
  \lleq{qac}
  p(\Lumtot \cond N, \bmphi, \bmy) &= \int \rmdx{\bmL} p(\Lumtot \cond \bmL, N, \bmphi, \bmy) p(\bmL \cond N, \bmphi, \bmy)\\
  \lleq{qaf}
  &= \int \rmdx{\bmL} \delta \left( \Lumtot - \sum_{j=1}^N \Lum_j  \right) \prod_{j=1}^N p(\Lum_j \cond \bmphi)\,,
\end{align}
hence $Q$ does not depend on $\bmy$ explicitly. In the next
subsections, we provide more details on our assumptions and provide
explicit expressions for \refeq{qaa}--\refeq{qaf}.

% As the reference frequency, we choose the bin center.

\subsection{Modeling the packet luminosity} \label{sec:pack-lum}

In general, the distribution $p(\Lum_j \cond \bmphi)$ could be a of a
different functional form in each bin. For simplicity, we assume it is
identical but may depend on parameters $\bmphi$ that need to be
inferred from $\bmy$.  To simplify the treatment, we assume that the
frequency bins are narrow or equivalently that $p(\Lumtot \cond \bmy)$
does not vary within a bin. This implies that every sample $\Lum_j$ in
the bin follows the same distribution.

What is the role of $\bmphi$? It contains all parameters needed to
model the entire spectrum that have to be learned from $\bmy$. To
begin, we include the Poisson means for each bin
\begin{align}
  \lleq{qad} \bmphi = \left( \bmlambda, \bmxi \right) = \left(
    {\lambda_1,\dotsc,\lambda_{n_b}}, \bmxi \right)\,.
\end{align}
\fred{How }
The number and the luminosity of the packets are assumed independent,
so the posterior factorizes as
\begin{align}
  \lleq{qadd}
  p(\bmphi \cond \bmy) = p(\bmlambda \cond \bmy) p(\bmxi \cond \bmy) \,.
\end{align}

Regarding the additional parameters $\bmxi$, we have essentially two
extreme options; both illustrated below in \refsec{example} and
\refsec{asymptotic}. The first option is to assume an explicit
\emph{parametric} form, for example some function $f$, such that
\begin{align}
  \lleq{qae}
  p(\Lum \cond \bmphi) = f(\Lum \cond \bmxi, \nu) \,,
\end{align}
where we now made the dependency on $\nu$ explicit again.  In this
way, we assume the same distribution for all frequencies and we only
need to learn the probable values of $\bmxi$ from the data $\bmy$,
$p(\bmxi \cond \bmy)$. The advantage is that $f(\Lum \cond \bmxi, \nu)$ may
be well constrained by $\bmy$ if $\npack$ is large and it is easy to
include further assumptions; e.g., $f$ should be continuous in $\nu$.

The second option is preferable if it is impossible to elicit an
appropriate explicit expression for $p(\Lum \cond \bmphi)$ \emph{and}
each bin has many packets; i.e., $n \gg 1$. Then we can include one or
more parameters for each bin, consider each bin independently, and
perform a \emph{nonparametric} analysis. Compared to the parametric
approach, we now typically have many more parameters, and each one is
determined only by the packets in one instead of all bins.

\subsection{Poisson analysis}

In order to constrain the Poisson mean $\lambda$ in a bin, we have one
measurement, the number of packets $n$ in that bin. Using Bayes' theorem, we find the posterior
\begin{align}
  \lleq{pmd}
  p(\lambda\cond n) %
  &= \frac{p(n\cond\lambda)\,p(\lambda)} {p(n)}
  \ =\ \frac{p(n\cond\lambda)\,p(\lambda)} %
  {\int \rmdx{\lambda} p(n\cond\lambda)\,p(\lambda)} \,.
\end{align}
% to find the posterior predictive of $N$ given $n$ taking into account
% all possible values of $\lambda$,
% \begin{align}
%   \lleq{pme}
%   p(N\cond n)
%   &= \int \rmdx{\lambda}p(N,\lambda\cond n)
%   \ =\ \int \rmdx{\lambda} p(N\cond\lambda, n)\,p(\lambda\cond n)
%   \ =\ \frac{\int \rmdx{\lambda} p(N\cond\lambda)\,p(n\cond\lambda)\,p(\lambda)}
%   {\int \rmdx{\lambda} p(n\cond\lambda)\,p(\lambda)}
% \end{align}
% where $p(N\cond\lambda,n) = p(N\cond\lambda)$ is independent of $n$.
Common uninformative priors are
\begin{align}
  \lleq{pmf}
  p(\lambda) = c \lambda^{-a},
\end{align}
where $a=0$ gives the uniform prior\footnote{For a uniform prior, the
  case $a{=}0$, we require of course a maximum $\lambda_{\rm max}$. As
  long as $\lambda_{\rm max}$ is large enough, the integral can
  effectively be taken to infinity.}, $a=1/2$ gives the Jeffreys and
reference prior, and $a=1$ is the transformation-group result
advocated by Jaynes~\cite[Ch. 12]{jaynes2003probability}. All choices
lead to an improper prior but the posterior is proper provided that
$n \ge 1$; i.e., at least one packet is observed in the bin. The
standard result for the posterior is
\begin{align}
  \lleq{pms}
  p(\lambda\cond n) &= \GammaDist(\lambda \cond \alpha = n-a+1, \beta = 1)\,,
\end{align}
where the  $\GammaDist$ distribution with shape parameter $\alpha$ and rate
parameter $\beta$ is defined as
\begin{align}
  \lleq{gma}
  \GammaDist(\lambda | \alpha, \beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} x^{\alpha-1} e^{-\beta x}\, , \; x > 0 .
\end{align}
% \begin{align}
%   \lleq{pmg}
%   p(n) &= %
%   \int_0^\infty d\lambda\, p(n\cond\lambda)\,p(\lambda)
%   \ =\
%   c \int_0^\infty d\lambda\, \frac{e^{-\lambda}\,\lambda^{n-a}}{n!}
%   \ =\ c\frac{(n-a)!}{n!} \, ,
%   \\
%   \lleq{pmh}
%   p(N\cond n) %
%   &= \frac{n!}{(n-a)!} \int_0^\infty d\lambda\,
%   \frac{e^{-2\lambda} \lambda^{N+n-a}}{N!\, n!}
%   \ =\ \frac{(N+n-a)!}{N!\,(n-a)!\,2^{N+n+1-a}} \, .
% \end{align}

In the case that there are many packets in every bin, some
simplifications arise as the exact distribution of individual packets
$p(\Lum \cond \bmphi, \nu)$ becomes irrelevant in the asymptotic
regime $n \to \infty$, which implies $\lambda \to \infty$. In that
case, the compound Poisson distribution can be approximated by a
Gaussian distribution provided that the second moment of
$\Lum, \expect{\Lum^2}$ is finite. For a compound Poisson process with Poisson mean $\lambda$,
the general result is that, even for $\lambda$ small,
\begin{align}
  \lleq{raa}
  \expect{Q} = \lambda \expect{L}, \qquad \variance{Q} = \lambda \expect{L^2} \,.
\end{align}

From Theorem 4.3.1 of \cite{bening2002generalized}, we can conclude
that the asymptotic distribution of $\Lumtot$ when
$\lambda \to \infty$, given the first and second moment of $\Lum$, is
\begin{align}
  \lleq{rab}
  P(\Lumtot \cond \bmphi) = \GaussianDist \left( \Lumtot \cond \expect{Q}, \variance{Q} \right) = \GaussianDist \left( \Lumtot \cond \lambda \expect{\Lum}, \lambda \expect{\Lum^2} \right).
\end{align}
This result is a drastic simplification as it eliminates both the sum over $N$ in \refeq{qab} and the integral over $\bmL$ in \refeq{qaf}.
That allows us to model $\Lumtot$ in each bin \emph{without} making concrete
assumptions about the functional form of $P(\Lum \cond \bmphi)$; i.e., we may now view $\bmphi$ as the collection of first and second moment in each bin in addition to $\lambda$. Continuing the notation of \refeq{qadd},
\begin{align}
  \lleq{rac}
  \bmphi =  \left( \bmlambda, \bmxi \right) = \left(\lambda_b, \expect{\Lum_b}, \expect{\Lum^2_b} \right)_{b=1}^B.
\end{align}
The price for this flexibility is that $\bmphi$ may now contain many
parameters, and each one is determined only from the samples in a
single bin. To obtain the posterior $p(\bmxi \cond \bmy)$, we may focus on
a single bin again
\begin{align}
  \lleq{rad}
  p(\bmxi \cond \bmy) = \prod_{b=1}^B p\left(\expect{\Lum_b}, \expect{\Lum^2_b} \cond \bmy\right)
\end{align}
and omit the subscript $b$. To learn the moments of $\Lum$, we follow
Jaynes~\cite[Ch. 7.10]{jaynes2003probability} and assign a Gaussian
sampling distribution
\begin{align}
  \lleq{rae}
  p(\Lum \cond \expect{\Lum}, \expect{\Lum^2} )
  &= \GaussianDist(\Lum \cond \mu = \expect{\Lum}, \sigma^2
  = \expect{\Lum^2} - \expect{\Lum}^2)
\end{align}
although the actual frequencies will be quite different from a
Gaussian. Our motivation is that this Gaussian represents our state of
knowledge \emph{before} observing $\bmy$ well. To make the
calculations easier, we assume a conjugate prior for unknown $\mu$ and
$\sigma^2$. The standard result for the posterior is
\begin{align}
  \lleq{raf}
  p(\mu, \sigma^2) &= \GaussianDist(\mu \cond \mu_0^\prime, \frac{\sigma^2}{n_0^\prime}) \InvGammaDist\left( \sigma^2 \cond \frac{\nu_0^\prime}{2}, \frac{\nu_0^\prime \sigma_0^{\prime 2}}{2} \right)
\end{align}
where the hyperparameters are updated from prior to posterior as
\begin{align}
  \lleq{rag}
  \expectest{\lum} &= \frac{1}{n} \sum_{i=1}^n \lum_i,\\
  \lleq{rah}  n_0^\prime &= n_0 + n,\\
  \lleq{rai}    \mu_0^\prime &= \frac{n_0 \mu_0 + n \expectest{\lum}}{n^\prime},\\
  \lleq{raj}    \nu_0^\prime &= \nu_0 + n,\\
  \lleq{rak}    \nu_0^\prime \sigma_0^{\prime 2} &= \nu_0 \sigma_0^{ 2} + \sum_{i=1}^n \left(\lum_j - \expectest{\lum}\right)^2 + \frac{n_0 n}{n^\prime} \left(\mu_0 - \expectest{\lum}\right)\,.
  \end{align}
  By default, we take a noninformative prior and set
  $n_0 = \mu_0 = \nu_0 = 0$, so the value $\sigma_0^2>0$ is
  irrelevant. The posterior
  $ p(\expect{\Lum}, \expect{\Lum^2} \cond \bmy)$ is given by
  \refeq{raf} with $\mu, \sigma^2$ replaced by
  $\expect{\Lum}, \expect{\Lum^2}$ as indicated in \refeq{rae} because
  the Jacobian determinant is one.

  Now all ingredients for the prediction of $\Lumtot$ in one bin are
  ready, and \refeq{qaa} becomes
\begin{align}
  \lleq{ral}
  p(\Lumtot \cond \bmy) &= \int \rmdx{\lambda} \rmdx{\expect{\Lum}} \rmdx{\expect{\Lum^2}} \GaussianDist \left( \Lumtot \cond \lambda \expect{\Lum}, \lambda \expect{\Lum^2} \right) p(\expect{\Lum}, \expect{\Lum^2} \cond \bmy) \,.
\end{align}
\fred{$\int \rmdx{\expect{\Lum}}$ looks ugly, I'd rather have other
  symbols, perhaps just use $\mu$ and $\sigma^2$. The transformation
  is one-to-one anyways.}
\fred{Could integrate directly, or use mgf to get moments of $  p(\Lumtot \cond \bmy)$. Then do mode finding, describe initial values and Laplace approximation. }

\end{document}
\subsection{Old}
%old text
The joint distribution then is
\begin{align}
  \lleq{cac}
  % Q, N \sim P(Q, N \cond \bmphi, )
  (N,\bmL) &\ \sim\ p(N,\bmL\cond \bmphi,\lambda, \nu)
  = p(\bmL\cond N,\bmphi, \nu)\; p(N\cond \lambda) = \prod_{j=1}^N p(\Lum_j\cond \bmphi, \nu) \; p(N\cond \lambda),
\end{align}
\fred{Do we need this equation?}
where we used that the Poisson term does not depend on $\nu$
explicitly. In general, the distribution of $\Lum_j$ could be a of a
different functional form in each bin. For simplicity, we assume it is
identical but may depend on parameters that differ from bin to bin. In
practice, the distribution of packets usually does not change abruptly
with frequency. We therefore assume that any parameter needed to fully
specify $p(\Lum_j)$ is given explicitly as a continuous function of $\nu$
and the unknown parameters $\bmphi$. This allows us to use a small
number of parameters to describe $p(\Lumtot)$ across the entire frequency
range instead of having a few parameters per bin which could lead to
tens of thousands of parameters for the applications we have in
mind. So we determine $\bmphi$ from fitting all simulated packets
simultaneously whereas $\lambda$ is a per-bin quantity.
\fred{Reformulate when asymptotic approx. included: it has thousands of parameters}

@@

We use Bayes' theorem to find the posterior distribution for
